Categorical Predictor Embeddings in Actuarial Modeling

### 2020-09-29 manuscript/manuscript.Rmd
add table
add table

A categorical variable $X_j, j \in J_{cat}$ takes as its value only one of a finite number of labels. Let the set of labels be $\mathcal{P}^j = \{p^j_1, p^j_2,\dots, p^j_{n_{\mathcal{P}^j}} \}$, where $n_\mathcal{P}^j = |\mathcal{P}^j|$ is the cardinality or number of levels, in $\mathcal{P}^j$. One-hot encoding maps each value $x_{i,j}$ of $X_j$ to $n_\mathcal{P}^j$ indicator variables, which take a value of $1$ if the label of $x_{i,j}$ corresponds to the level of the indicator variable, and $0$ otherwise. One-hot encoding is often used in the machine learning community while the statistical community often favors dummy coding, which, instead of assigning $n_p^j$ indicator variables, assigns one of the levels of the categories as a baseline, and maps all of the other $n_p^j -1$ variables to indicator variables.

@@

| Variable                                 | Type                                     |
|------------------------------------------|------------------------------------------|
| Building insurance coverage              | Numeric                                  |
| Basement enclosure type                  | Categorical                              |
| Number of floors in the insured building | Categorical (binned in original dataset) |
| Flood zone                               | Categorical (high cardinality)           |
| Primary residence                        | Categorical (binary indicator)           |

>>>>

A categorical variable $X_j, j \in J_{cat}$ takes as its value only one of a finite number of labels. Let the set of labels be $\mathcal{P}^j = \{p^j_1, p^j_2,\dots, p^j_{n_{\mathcal{P}^j}} \}$, where $n_\mathcal{P}^j = |\mathcal{P}^j|$ is the cardinality or number of levels, in $\mathcal{P}^j$. One-hot encoding maps each value $x_{i,j}$ of $X_j$ to $n_\mathcal{P}^j$ indicator variables, which take a value of $1$ if the label of $x_{i,j}$ corresponds to the level of the indicator variable, and $0$ otherwise. An example of one-hot encoding is shown in Table \ref{tab:onehot}.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
require(dplyr)
require(cellar)
require(recipes)
require(data.table)

# nfip = cellar_pull("nfip_claims", vintage = NULL)
# 
# states = nfip %>% select(state) %>% unique()
# 
# states %>% fwrite("manuscript/states.csv")

states= fread("../manuscript/states.csv")

sample_dat = sample_n(states,5) %>% arrange(state)

rec = recipe(~state, data = sample_dat)

dummies <- rec %>% step_dummy(state)
dummies <- prep(dummies, training = sample_dat)

dummy_data <- bake(dummies, new_data =sample_dat)

cbind(sample_dat, dummy_data) %>% knitr::kable(label = "tab:onehot", caption = "Example one-hot encoding of the state variable")
```

One-hot encoding is often used in the machine learning community while the statistical community often favors dummy coding, which, instead of assigning $n_p^j$ indicator variables, assigns one of the levels of the categories as a baseline, and maps all of the other $n_p^j -1$ variables to indicator variables.

@@

+------------------------------------------+------------------------------------------+
| Variable                                 | Type                                     |
+==========================================+==========================================+
| Building insurance coverage              | Numeric                                  |
+------------------------------------------+------------------------------------------+
| Basement enclosure type                  | Categorical                              |
+------------------------------------------+------------------------------------------+
| Number of floors in the insured building | Categorical (binned in original dataset) |
+------------------------------------------+------------------------------------------+
| Flood zone                               | Categorical (high cardinality)           |
+------------------------------------------+------------------------------------------+
| Primary residence                        | Categorical (binary indicator)           |
+------------------------------------------+------------------------------------------+
